Smart Model Elimination for Efficient Automated Machine Learning in Healthcare

### 2024-06-27 paper/paper.tex
some tense stuff

Through this survey, we uncovered a key continuity in these frameworks' design: frameworks must run all of their models to maximize accuracy at the expense of efficiency. Given the specific nature of certain tabular data, there are some models that are less likely to perform well, thus running and training these models is wasteful. Cutting these models completely, however, is not optimal, as they are the best option in a minority of cases. We hypothesized that using a combination of multi-layer machine learning implementation and meta-feature extraction would yield an equally versatile system that minimizes time spent on training without sacrificing accuracy.

@@

We considered 28 different models, including all supervised and semi-supervised sklearn [12] classification models and common boosting models such as XGBoost [13], LightGBM [14], and Catboost [15]. The majority of these models are implemented in the frameworks mentioned in Table 1.

To implement the model elimination process, we train a multilabel-regression boosting model called the SME Model-Ranker on 268 kaggle datasets. We choose medical datasets and focus our approach on binary classification. We extract the attributes in Table 2 from each dataset: which are the inputs for the boosting model. We then run a "dumb", brute-force framework that trains every model every each dataset. The accuracy of every model is ranked, and every model is treated as a label with its value being its rank. We train the SME to predict the rank of each model, given the extracted attributes. SME produces a median spearman correlation of 0.7301 between the predicted rank and the actual rank and a median p-value of 1.032e-05. Furthermore, we guarantee that we can predict a top two model within the predicted top eight 90.12\% of the time.

@@

\section{Evaluation}
\subsection{Experiment Design}
To evaluate our framework, we choose six different datasets of diverse size. We ran SMEML on each dataset for 20 iterations of tuning. We compared this against a "dumb" mode which simply runs every model on the dataset. We used the accuracy and time to train as our metrics. All experiments are run on an Ubuntu server with Intel Xeon E5520 @ 2.27GHz and 24GB of RAM.

@@

The first observation made was that accuracy did not change for 4 of the 6 benchmarks. The other two benchmarks on eye states and diabetes prediction represented a loss of accuracy for SMEML by 0.13\% and 1.95\%, respectively. Training time, however, favored SMEML in every trial. SMEML outperformed the dummy framework by 173.96 seconds in its worst performance on the lung cancer trial and by 383.94 seconds in its best performance on the eye state dataset. Relative to the dummy's performance, a percent decrease in time to run for SMEML of at least 46.78\% could be seen, with a decrease of up to 69.22\% observerd as well. SMEML's worst performance by percent decrease was also the performance in which it returned perfect accuracy. 
\section{Discussion}
\subsection{Summary}
SMEML consistently achieves comparable accuracy to its dumb counterpart while significantly reducing the time required to train the model. Notably, SMEML's performance is not limited by the size of the dataset, achieving its best runtimes--first, third, and fourth fastest overall--on 'large' datasets (those exceeding 1000 rows). SMEML also chose less common algorithms, such as selecting SGDClassifier for the Pima Indians Diabetes Dataset [19] rather than defaulting to industry-standard boosting algorithms [23]. This efficiency and versatility is important in third world medicine where there is a constant influx of new patient data that must be processed quickly. A practical demonstration of this software's real-world application can be found on our SmartDx website [24]. Source code can be found at [25].
\subsection{Limitations}
Our implementation of SMEML has limitations. These include a small training set due to lack of available datasets, finite computing power, time constraints, and a lack of access to paticular scientific literature. Our framework only improves on the model selection process and does not address the various other aspects of AutoML. Therefore, SMEML is not comparable to other full-featured AutoML frameworks.
\subsection{Future Work}
Future work will include increasing the size of the training set, implementing a fully fledged AutoML framework, and conducting a detailed experiment comparing SMEML to other popular AutoML frameworks.

>>>>

Through this survey, we uncover a key continuity in these frameworks' design: frameworks must run all of their models to maximize accuracy at the expense of efficiency. Given the specific nature of certain tabular data, there are some models that are less likely to perform well, thus running and training these models is wasteful. Cutting these models completely, however, is not optimal, as they are the best option in a minority of cases. We hypothesize that using a combination of multi-layer machine learning implementation and meta-feature extraction would yield an equally versatile system that minimizes time spent on training without sacrificing accuracy.

@@

We consider 28 different models, including all supervised and semi-supervised sklearn [12] classification models and common boosting models such as XGBoost [13], LightGBM [14], and Catboost [15]. The majority of these models are implemented in the frameworks mentioned in Table 1.

To implement the model elimination process, we train a multilabel-regression boosting model called the SME Model-Ranker on 268 kaggle datasets. We choose medical datasets and focus our approach on binary classification. We extract the attributes in Table 2 from each dataset, which are the inputs for the boosting model. Then a "dumb", brute-force framework that trains every model on every dataset is run. The accuracy of every model is ranked, and every model is treated as a label with its value being its rank. We train the SME to predict the rank of each model, given the extracted attributes. SME produces a median spearman correlation of 0.7301 between the predicted rank and the actual rank and a median p-value of 1.032e-05. Furthermore, we guarantee that we can predict a top two model within the predicted top eight 90.12\% of the time.

@@

\section{Evaluation}
\subsection{Experiment Design}
To evaluate our framework, we choose six different datasets of diverse size. We run SMEML on each dataset for 20 iterations of tuning. We compare this against a "dumb" mode which simply runs every model on the dataset. We use the accuracy and time to train as our metrics. All experiments are run on an Ubuntu server with Intel Xeon E5520 @ 2.27GHz and 24GB of RAM.

@@

The first observation made was that accuracy did not change for 4 of the 6 benchmarks. The other two benchmarks on eye states and diabetes prediction represented a loss of accuracy for SMEML by 0.13\% and 1.95\%, respectively. Training time, however, favores SMEML in every trial. SMEML outperforms the dummy framework by 173.96 seconds in its worst performance on the lung cancer trial and by 383.94 seconds in its best performance on the eye state dataset. Relative to the dummy's performance, a percent decrease in time to run for SMEML of at least 46.78\% could be seen, with a decrease of up to 69.22\% observerd as well. SMEML's worst performance by percent decrease is also the performance in which it returned perfect accuracy. 
\section{Discussion}
\subsection{Summary}
SMEML consistently achieves comparable accuracy to its dumb counterpart while significantly reducing the time required to train the model. Notably, SMEML's performance is not limited by the size of the dataset, achieving its best runtimes--first, third, and fourth fastest overall--on 'large' datasets (those exceeding 1000 rows). SMEML also chose less common algorithms, such as selecting SGDClassifier for the Pima Indians Diabetes Dataset [19] rather than defaulting to industry-standard boosting algorithms [23]. This efficiency and versatility is important in third world medicine where there is a constant influx of new patient data that must be processed quickly. A practical demonstration of this software's real-world application can be found on our SmartDx website [24]. Source code can be found at [25].
\subsection{Limitations}
Our implementation of SMEML has limitations. These include a small training set due to lack of available datasets, finite computing power, time constraints, and a lack of access to paticular scientific literature. Our framework only improves on the model selection process and does not address the various other aspects of AutoML. Therefore, SMEML is not comparable to other full-featured AutoML frameworks.
\subsection{Future Work}
Future work will include increasing the size of the training set, implementing a fully fledged AutoML framework, and conducting a detailed experiment comparing SMEML to other popular AutoML frameworks.
